Solving an optimization problem employing restraint programming

### 2015-12-13 proj2/doc/article.tex
Relatorio acabado

% This is LLNCS.DEM the demonstration file of
% the LaTeX macro package from Springer-Verlag
% for Lecture Notes in Computer Science,
% version 2.4 for LaTeX2e as of 16. April 2010
%
\documentclass{llncs}
%
\usepackage{makeidx}  % allows for indexgeneration
\usepackage[utf8]{inputenc}
%
\begin{document}
\tableofcontents
%
%\mainmatter              % start of the contributions
%
\title{Solving an optimization problem employing restraint programming}
%
\titlerunning{Restraint Programming}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{João Cabral\inst{1} \and João Mota\inst{2}}
%
\authorrunning{João Cabral \and João Mota} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
\tocauthor{João Cabral, João Mota}
%
\institute{Faculdade de Engenharia da Universidade do Porto, Porto, Rua Roberto Frias s/n 4200-465, Portugal,\\
\email{(up201303462, up201304395)@fe.up.pt}}

@@

\keywords{optimization, restraint programming, PROLOG}
\end{abstract}
%
ISTO TA A FICAR APERTADO D PAGS, SE EXCEDER TIRA-SE OS VERBATIMS TODOS E OS PROFS PODEM IR LER AOS ANEXOS Q SE FODEM
\section{Introduction}
Within the context of the Logic Programming course of MIEIC, the authors of this paper were asked to solve an optimization problem in the most efficient way possible using the restraint programming capabilities of SICStus PROLOG. The optimization problem under study is that of producing a working schedule for a language school that maximizes it's (monetary) gains,  given a number of inputs of variable length, aswell as a number of restrictions of various types. The rest of the paper is structured as follows:

@@

\subsection{Restraints}
% FALTA BOTAR AQUI CENAS
The restraints for this particular problem are listed in section 2. As such this section will focus mainly on restrictions' implementation aspects. All restricitions concerning monetary values are approached during the profit calculating phase of the solution, which is further explained in the next section of this paper. For the restriction concerning the number of hours of class a certain course can have per week, the program will gather the values for the hours taught each day and say that their sum must be grater than or equal to 2 and lesser than or equal to 4.

@@

@@

\begin{verbatim}
labeling([ff,enum,down,maximize(Profit)], Flat).
\end{verbatim}

\subsection{Search Strategy}
 NUM SEI OQ HEI-DE POR AQUI

@@

 \section{Results}
 %BOTA GRAFICOS E TAL

 \section{Conclusions and Future Work}
 Having implemented a solution to the problem under study, the paper's authors are able to appreciate the advantages of employing restraint programming in the solving of problems that are similar to the one studied here. Indeed, the implemented solution shows that the use of this technology allows for a relatively simple implementation of a solution, when compared to possible implementations using other technologies.
%aspetos a melhorar???

>>>>

% This is LLNCS.DEM the demonstration file of
% the LaTeX macro package from Springer-Verlag
% for Lecture Notes in Computer Science,
% version 2.4 for LaTeX2e as of 16. April 2010
%
\documentclass{llncs}
%
\usepackage{makeidx}  % allows for indexgeneration
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{float}
\usepackage{placeins}
%
\begin{document}
\tableofcontents
%
%\mainmatter              % start of the contributions
%
\title{Solving an optimization problem employing restraint programming}
%
\titlerunning{Restraint Programming}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{João Cabral\inst{1} \and João Mota\inst{2}}
%
\authorrunning{João Cabral \and João Mota} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
\tocauthor{João Cabral, João Mota}
%
\institute{Faculdade de Engenharia da Universidade do Porto, Porto, Rua Roberto Frias s/n 4200-465, Portugal,\\
\email{(up201303462, up201304395)@fe.up.pt}}

@@

\keywords{optimization, restraint programming, PROLOG}
\end{abstract}
%
%ISTO TA A FICAR APERTADO D PAGS, SE EXCEDER TIRA-SE OS VERBATIMS TODOS E OS PROFS PODEM IR LER AOS ANEXOS Q SE FODEM
\section{Introduction}
Within the context of the Logic Programming course of MIEIC, the authors of this paper were asked to solve an optimization problem in the most efficient way possible using the restraint programming capabilities of SICStus PROLOG. The optimization problem under study is that of producing a working schedule for a language school that maximizes it's (monetary) gains,  given a number of inputs of variable length, aswell as a number of restrictions of various types. The rest of the paper is structured as follows:

@@

\subsection{Constraints}
% FALTA BOTAR AQUI CENAS
The constraints for this particular problem are listed in section 2. As such this section will focus mainly on constraints' implementation aspects. All constraints concerning monetary values are approached during the profit calculating phase of the solution, which is further explained in the next section of this paper. For the constraint concerning the number of hours of class a certain course can have per week, the program will gather the values for the hours taught each day and say that their sum must be grater than or equal to 2 and lesser than or equal to 4.

@@

In the interest of reducing the domain and the search space to feasible levels the authors implemented a further constraint that limits each class to only having a 2 hour duration. This means that each course can have either one 2 hour class or 2 2 hour classes.

@@

\begin{verbatim}
labeling([ff,leftmost,down,maximize(Profit)], Flat).
\end{verbatim}

\subsection{Search Strategy}
Establishing an optimal search strategy for this problem presents difficulties due to the large variety of valid inputs possibly benefitting from a variety of different options. The search strategies selected are presented with a brief rationale for their choice. The tests run on each set of options are show on table \ref{table:speedup}.

For choice making the \textit{step} option was selected over the common \textit{bisect} option. It allows for a speedier pruning of the domain space because, in conjunction with the \textit{down} option it leads to higher profit solutions as the first solutions. The discovery of high profit solutions early in the branch-and-bound algorithm allows for larger prunings to be made on the search tree. This results in marginally faster times when compared to \textit{bisect}.

Given the goals of maximizing the cost being closely tied with maximizing the number of hours taught at the school the obvious choice of searching though domains in descendant order with the option \textit{down} was made. This resulted in significant speedups when compared to the default option \textit{up}.

\begin{table}[]
\caption{6 professors, 5 languages, 4 classrooms test}
\label{table:speedup}
\centering
\begin{tabular}{l|l|l|}
\cline{2-3}
                                      & time(s) & speedup \\ \hline
\multicolumn{1}{|l|}{up,bisect,ff}    & 520.052 & 1x      \\ \hline
\multicolumn{1}{|l|}{down,bisect,ff,} & 20.725  & 25.09x  \\ \hline
\multicolumn{1}{|l|}{down,step,ff}  & 21.349  & 25.55x  \\ \hline
\end{tabular}

\end{table}

@@

 \section{Results}
The following graphics plot the performance results obtained with the modeling and constraints chosen, when increasing the number of classrooms, professors and languages being taught. All other variables are not changing in each series, to provide a accurate representation of the performance of the program.

\begin{figure}[h]
  \centering
    \includegraphics[width=1\textwidth]{professor.png}
\end{figure}

\begin{figure}[h]
  \centering
    \includegraphics[width=1\textwidth]{classrooms.png}
\end{figure}

\begin{figure}[h]
  \centering
    \includegraphics[width=1\textwidth]{language.png}
\end{figure}

\FloatBarrier
 \section{Conclusions and Future Work}
 Having implemented a solution to the problem under study, the paper's authors are able to appreciate the advantages of employing restraint programming in the solving of problems that are similar to the one studied here. Indeed, the implemented solution shows that the use of this technology allows for a relatively simple implementation of a solution, when compared to possible implementations using other technologies.

It is possible to observe the effect on performance of the establishment of constraints, which works somewhat conterintuitively in that the imposition of further constraints can be extremely beneficial for performance, achieving sometimes surprising speedups by imposing constraints. One should however note that some constraints, particularly those that are not global, can also significantly degrade performance. The use of reification in particular, which the authors attempted in one of the first implementations, proved to be too costly despite the advantages in code simplicity it provided. It was thus replaced with a global constraint.

The solution now presented suffers from a significant degradation in performance as the problem size expands. This is partly due to the intrinsic characteristics of the problem that mandates the use of multiple decision variables with  relatively large domains, compounded by the fact that each decision made by the solver can have a relatively modest effect on the size of the remaining domains when it propagates. However, the modelization chosen is also not optimal as it allows, within the constraints of the problem, for multiple, equivalent, solutions that while valid and distinct do not represent any sort of improvement or even to relevant alternatives. For example, professors who can give classes in a large range of days can provoke an exponential growth of the solution space, even when constrained to the optimal solutions, which the solver is forced to explore. Relevant future work includes improvements to the modelization of the problem and possibly the implementation of a heuristic of selection of domain variables to allow for quicker constraint of the domains and possibly discarding exploration of useless, equivalent paths in the search tree.
%aspetos a melhorar???
